---
title: "Mini-Project 4"
author: "Emily Liu"
date: "April 28, 2018"
output: html_document
---
```{r}
library(mdsr)
library(RMySQL)

library(leaflet)
library(sf)

library(tidyverse)

db <- dbConnect_scidb(dbname = "citibike")

```

```{r}
test_set <- db %>%
          dbGetQuery("SELECT * 
             FROM trips
             LIMIT 100;")
test_set

test_stations <- db %>%
  dbGetQuery("SELECT *
             FROM station_months;")

test_stations %>%
  distinct(station_id) %>%
  arrange(station_id)


test_stations_sum <- db %>%
  dbGetQuery("SELECT *
             FROM station_summary;")
```

```{r}
# Top 10 stations overall
test_set %>%
  arrange(start_station_id)
test_set %>% 
  select(start_station_id) %>%
  group_by(start_station_id) %>%
  summarize(N = n()) %>%
  arrange(desc(N)) %>%
  head(10)
  
```

```{r}
# Top 10 stations for women
women_10 <- test_set %>% 
  filter(gender == 2) %>%
  select(start_station_id) %>%
  group_by(start_station_id) %>%
  summarize(N = n()) %>%
  arrange(desc(N)) %>%
  head(10)

# Top 10 stations for subscribers
test_set %>% 
  filter(user_type == 'Subscriber') %>%
  select(start_station_id) %>%
  group_by(start_station_id) %>%
  summarize(N = n()) %>%
  arrange(desc(N)) %>%
  head(10)
```

```{r}
top_10_women <- women_10 %>%
  left_join(test_stations_sum, by = c("start_station_id" = "station_id"))

View(top_10_women)
```

```{r}
#finding top 10 start stations overall
db %>%
  dbGetQuery("
SELECT trips.start_station_id,  ss.name, ss.lat, ss.lon, sum(1) AS N
FROM citibike.trips
JOIN citibike.station_summary ss ON trips.start_station_id = ss.station_id
GROUP BY trips.start_station_id
ORDER BY N desc
LIMIT 10;")

#finding top 10 start stations visited by women
db %>%
  dbGetQuery("
SELECT trips.start_station_id,  ss.name, ss.lat, ss.lon, sum(1) AS N
FROM citibike.trips
JOIN citibike.station_summary ss ON trips.start_station_id = ss.station_id
WHERE gender = 2
GROUP BY trips.start_station_id
ORDER BY N desc
LIMIT 10;")
```
```{r}
# Points for overall start stations
all_start_pts <- tribble (
  ~name, ~lat, ~lon,
  "Pershing Square North", -73.97771, 40.75187,
  "E 17 St & Broadway", -73.99009, 40.73705,
  "Broadway & E 22 St", -73.98955, 40.74034,
  "W 21 St & 6 Ave", -73.99416, 40.74174,
  "West St & Chambers St", -74.01322, 40.71755,
  "8 Ave & W 31 St", -73.99468, 40.75059,
  "8 Ave & W 33 St", -73.99393, 40.75155,
  "W 20 St & 11 Ave", -74.00776, 40.74674,
  "12 Ave & W 40 St", -74.00278, 40.76088,
  "W 41 St & 8 Ave", -73.99003, 40.75461
)

all_start_sf <- all_start_pts %>%
  st_as_sf(coords = c("lat", "lon")) %>%
  st_set_crs(4326)

leaflet() %>%
  addTiles() %>%
  addMarkers(data = all_start_sf, popup = ~name)
```


```{r}
#finding top 10 most visited stations for everyone
db %>%
  dbGetQuery("
SELECT *, num_starts + num_stops AS visits
FROM citibike.station_summary ss
ORDER BY visits desc
LIMIT 10;")
```

```{r}
#finding most visited stations for females

#start stations and their number of visits for females
start_tally <- db %>%
  dbGetQuery("
SELECT trips.start_station_id, sum(1) AS N_start
FROM citibike.trips
WHERE gender = 2
GROUP BY trips.start_station_id
ORDER BY N_start desc;")

end_tally <- db %>%
  dbGetQuery("
SELECT trips.end_station_id, sum(1) AS N_end
FROM citibike.trips
WHERE gender = 2
GROUP BY trips.end_station_id
ORDER BY N_end desc;")

#trying to full join start_tally and end_tally in order to calculate the total number of visits, but need to have same column name? 
rename(start_tally, "start_station_id" = "station_id")
rename(end_tally, "end_station_id" = "station_id")

female_visits <- full_join(start_tally, end_tally, by = "station_id"))
